Marine Conservation and Fishing Effort Markets

### 2019-05-04 docs/slides/NCEAS_presentation.Rmd
Updatesd slides for NCEAS

---
title: "Marine protected areas under effort-control management"
author:
  - Villaseñor-Derbez J.C.$^1$
  - Lynham J.$^2$
  - Costello C.$^{1, 3}$
institute: 
   - $^1$Bren School, UCSB
   - $^2$Department of Economics, UHM
   - $^3$emLAB, UCSB
subtitle: "NCEAS-Future Earth Working Group"
date: "May 6, 2019"
output: beamer_presentation
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

@@

@@

- Approach: model and assumptions

- Data

@@

## Background

```{r}

suppressPackageStartupMessages({
  library(startR)
  library(here)
  library(sf)
  library(tidyverse)
  library(ggdark)
})

source(here("scripts", "functions", "st_rotate.R"))

rast <- readRDS(here("data", "rasterized_effort_by_region.rds")) %>% 
  mutate(post = fct_relevel(
    ifelse(year < 2015, "Before", "After"),
    "Before")) %>% 
  group_by(post, x, y) %>%
  summarize(days = log10(sum(hours) / 24)) %>%
  ungroup() %>% 
  filter(between(x, 160, 210),
         between(y, -15, 10))

pipa <- read_sf(dsn = here("data", "spatial", "PIPA"),
                layer = "PIPA") %>% 
  st_rotate()

pna <- read_sf(dsn = here("data", "spatial", "PNA_EEZ"),
               layer = "PNA_EEZ") %>% 
  select(ISO_Ter3) %>% 
  st_simplify(11) %>% 
  st_rotate() %>% 
  st_union(by_feature = T) %>% 
  st_crop(y = c(xmin = 160, xmax = 210, ymin = -15, ymax = 10))

colors = c(
  "#0C276C",
  "#0B3B74",
  "#0A4E7C",
  "#096284",
  "#08768C",
  "#078994",
  "#059D9B",
  "#04B0A3",
  "#03C4AB",
  "#02D8B3",
  "#01EBBB",
  "#00FFC3"
)

ggplot() +
  geom_raster(data = rast,
              mapping = aes(x = x, y = y, fill = days)) +
  geom_sf(data = pna, fill = "transparent", color = "steelblue") +
  geom_sf(data = pipa, fill = "transparent", color = "red", size = 1) +
  facet_wrap(~post, ncol = 1) +
  # scale_fill_viridis_c() +
  scale_fill_gradientn(colors = colors) +
  guides(fill = guide_colorbar(title = quo(log[10]~"(days)"),
                               frame.colour = "white",
                               ticks.colour = "white")) +
  dark_theme_minimal() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(strip.text = element_text(size = 12)) +
  labs(x = "", y = "") +
  ggtitle(label = "Phoenix Islands Protected Area (PIPA), Kiribati")
```

## Scope and Motivation

@@

  - How can these costs be reduced or eliminated?

@@

- 10-patch model of the fishery (9 patches under VDS, one patch under Open Access)

- Vessel-level profits for fishing in patch $i$ are:

$$
\Pi_i(E_i,X_i,R) = pqE_iX_i\Omega_i-cE_i^\beta
$$

- In patch 1, $\Omega_1 = \theta + (1 - \theta)(1 - R)$

- $\Omega = 1$ for patches $i > 1$

@@

$$
\pi_i(E_i) = pqX_i\Omega_i- \beta cE_i^{\beta-1}
$$

@@

@@

@@

```{r}
knitr::include_graphics(path = here("docs", "slides", "img", "demand_curves.png"))
```

## Less movement increases costs

@@

## Vessel-day trading reduces costs

@@

## What can we learn from PIPA

@@

- 397,447 Km^2^

- 1.5 times area of Ecuador

- ~2% of total PNA area

- ~10% of fishing activity in PNA

## Generalizability to Palau

>>>>

---
title: "Marine Conservation and Fishing Effort Markets"
author:
  - Villaseñor-Derbez J.C.$^1$
  - Lynham J.$^2$
  - Costello C.$^{1, 3}$
institute: 
   - $^1$Bren School, UCSB
   - $^2$Department of Economics, UHM
   - $^3$emLAB, UCSB
subtitle: "NCEAS-Future Earth Working Group"
date: "May 6, 2019"
output: 
  beamer_presentation: 
    theme: Darmstadt
header-includes:
  - \usepackage{media9}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE#,
                      # fig.align = "center"
                      )

suppressPackageStartupMessages({
  library(here)
})
```

@@

- Background and motivation

@@

- Approach

  - Model and assumptions

  - Phoenix Island Protected Area

@@

# Background and motivation

## Motivation

\begin{center}
	\includemedia[
    label=video1,
    width=300pt,height=225pt,%width=300pt,height=225pt,
	]{}{GFW_video.swf}

    \mediabutton[mediacommand=video1:playPause,
    overface=\color{blue}{\fbox{\strut Play/Pause}},
    downface=\color{red}{\fbox{\strut Play/Pause}}
    ]{\fbox{\strut Play/Pause}}
\end{center}

## Motivation

### Relationship between closure size and losses

Does a 30\% closure mean a 30\% loss in revenue?

# Scope and Approach

## Scope

Can economic instruments incentivize conservation?

@@

  - Can these costs be reduced or eliminated?

@@

- 10-patch model of the tuna purse-seine fishery

  - 9 patches under VDS
  - one patch under Open Access

- Profits from fishing in patch $i$ are:

$$\Pi_i(E_i,X_i,R) = pqE_iX_i\Omega_i-cE_i^\beta$$

$$\Omega_1 = \theta + (1 - \theta)(1 - R)$$

@@

$$\pi_i(E_i) = pqX_i\Omega_i- \beta cE_i^{\beta-1}$$

@@

- Patch-level effort is then

$$E_i = \left(\frac{pqX_i\Omega_i - \pi_i}{\beta c }\right) ^ { \frac{1}{\beta - 1}}$$

@@

$$\bar{E} = \sum_{i = 1}^9\left(\frac{pqX_i\Omega_i - \pi}{\beta c }\right) ^ {\frac{1}{\beta - 1}}$$

@@

```{r}
# Recortar para que salgan hasta 3000 y 10000 solamente
knitr::include_graphics(path = here("docs", "slides", "img", "demand_curves.png"))
```

# Model predictions

## Model predictions

### Relationship between closure size and losses

Does a 30\% closure mean a 30\% loss in revenue?

## Losses with no trading

@@

## Model predictions

### A market for fishing efforts

- What is the effect of trading?

- Can costs be reduced?

## Trading significantly reduces costs

```{r}
knitr::include_graphics(here("docs", "slides", "img", "cost_with_trading_long_axis.png"))
```

## Trading significantly reduces costs

@@

- Vessel-days are allocated to countries each year

- Combination of biomass and in-EEZ usage

$$
E_i^* = \alpha \left(\frac{\sum_{\tau = 0}^{\hat{\tau}}E_{i,t-\tau}}{\bar{E}\hat{\tau}} \right) +
(1 - \alpha) \left(\frac{\sum_{\tau = 0}^{\hat{\tau}}X_{i,t-\tau}}{\bar{X}\hat{\tau}} \right)
$$

- $\alpha$ is a weight on historical effort ($E_i$) and historical biomass ($B_i$)

What is the importance of allocation rules?

## "*Biomass-based*" allocation reduces costs

```{r}
knitr::include_graphics(here("docs", "slides", "img", "allocation_cost_plot.pdf"))
```

# PIPA

## What can we learn from PIPA?

PIPA:

@@

- 397,447 $Km^2$

- ~2\% of total PNA area

- ~10\% of fishing activity in PNA

GFW data:

- 313 tuna purse seine vessels that fished in PNA

- 92 Fished both before and after PIPA implementation:

  - 64 Fished within PIPA before implementation

- Devideded into three groups

## Two fleets

```{r}
knitr::include_graphics(here("docs", "slides", "img", "sample_tracks.png"))
```

## Effort displacement

```{r}
knitr::include_graphics(here("docs", "img", "all_PS_VDS_year.pdf"))
```

## Effort displacement in Kiribati

```{r}
knitr::include_graphics(here("docs", "img", "all_PS_VDS_KIR_year.pdf"))
```

## Effort displacement raster

```{r}
knitr::include_graphics(here("docs", "img", "rasterized_change.png"))
```

# Generalizability to Palau

# Extra slides - GFW data

## Effort vs catches

```{r}
knitr::include_graphics(here("docs", "slides", "img", "catch_effort_plot.png"))
```

## Effort and revenues

```{r}
knitr::include_graphics(here("docs", "img", "revenue_FFA_GFW_linear.pdf"))
```

## Vessel statistics

A short table with information on number of vessels per year and class

Palau-specific information about longliners
